---
title: "Recommandation System"
output: 
  html_notebook:
    toc: yes
    toc_float: yes
    toc_depth: 4
author: Urvi Chandreshkumar Sheth
---

#Problem 2
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library('textreuse')
library(magrittr)
library(dplyr) 
library(stats)
library(data.table)
library(reshape2)
library(lsa)
```


```{r}
rm(list=ls())
userid <-  20430232%%671
movies.file <- read.csv(file=file.path("ml-latest-small", "movies.csv"), header = T, sep = ",")
ratings.file <- read.csv(file=file.path("ml-latest-small", "ratings.csv"), header = T, sep = ",")
```

###User Profile

### get movies rated for user_id from ratings.file 
### get details of movie rated by user from movies.csv 
```{r}
user.movies <- ratings.file[which(ratings.file$userId == userid), ]
user.genere.list <- movies.file[which(movies.file$movieId %in% user.movies$movieId),]
user.genere.list$genres <- as.character(user.genere.list$genres)
row.names(user.genere.list) <- 1:nrow(user.genere.list)
```

### now take each genere from one entry and put it into different dimensions 
### create matrix of howmany number of movies watched by user by genres 
### store it in the file in ml-latest-small 
```{r}
allGenere <- c("Action", "Adventure", "Animation", "Children", "Comedy", "Crime", "Documentary", "Drama", "Fantasy","Film-Noir", "Horror", "IMAX", "Musical", "Mystery", "Romance", "Sci-Fi", "Thriller", "War", "Western", "(no genres listed)")

newrowcount <- nrow(user.genere.list) + 1 # extra row for avg

profile.user <- matrix(as.numeric(0),nrow = newrowcount,ncol = length(allGenere))

colnames(profile.user) <- factor(allGenere)

rownamesnew <- user.genere.list$movieId 
rownamesnew[length(rownamesnew )+ 1] <- "AVG"   #adding avg row
row.names(profile.user) <- rownamesnew

profile.user[nrow(user.genere.list),1] <- 0


for (i in 1:nrow(user.genere.list)) {
    list.genre <- unlist(strsplit(user.genere.list[i,]$genres,split="[|]"))
    for (k in 1:length(list.genre)){
        profile.user[i,list.genre[k]] <- as.integer(1)
    }
}
#calculating mean
for(j in 1:length(allGenere)){
  profile.user[newrowcount,j] <- mean(as.numeric(profile.user[1:nrow(user.genere.list),j]))
}

head(profile.user)
filename <- paste("userprofile_",as.character(userid),".csv", sep = "")
write.csv(profile.user,file =file.path("ml-latest-small", filename))

```


### Movie Profile
### sample 10 movies
```{r}
sample_length <-  10
random_sample <- movies.file[sample(1:nrow(movies.file), sample_length, replace = FALSE), ]
random_sample$genres <- as.character(random_sample$genres)
rownames(random_sample) <- 1:sample_length
random_sample
```

### create movie profile
```{r}
profile.movie <- matrix(as.numeric(0),nrow = sample_length ,ncol = length(allGenere))
colnames(profile.movie) <- factor(allGenere)
row.names(profile.movie) <- random_sample$movieId

for (i in 1:nrow(random_sample)) {
    list.genre <- unlist(strsplit(random_sample[i,]$genres,split="[|]"))
    for (k in 1:length(list.genre)){
        profile.movie[i,list.genre[k]] <- as.integer(1)
    }
}

head(profile.movie)
filename <- paste("movieprofile_",as.character(userid),".csv", sep = "")
write.csv(profile.movie,file =file.path("ml-latest-small", filename))
```

### cosine sim
```{r}
my.cosine <- function(x, y) {
  sum(x*y)/(norm(x, type="2") * norm(y, type="2"))
}

final_random_sample <- random_sample[c(-3)]
Avg_vector<- profile.user[nrow(profile.user), ] 

for(i in 1:nrow(profile.movie)){
    final_random_sample[i,"similarity"] <- my.cosine(profile.movie[i,], Avg_vector)    
}
final_random_sample <- final_random_sample[order(final_random_sample$similarity,decreasing = TRUE),]
final_top_5 <- final_random_sample[1:5,]
rownames(final_top_5) <- NULL

final_top_5
```

###Collaborative Filtering

```{r}
rm(list=ls())
userid <-  191
movies.file <- read.csv(file=file.path("ml-latest-small", "movies.csv"), header = T, sep = ",")
ratings.file <- read.csv(file=file.path("ml-latest-small", "ratings.csv"), header = T, sep = ",")

movie.by.user <- ratings.file %>%
  left_join(movies.file, by = c("movieId" = "movieId")) %>%
  filter(ratings.file$userId == userid)

movie.by.user
```

### test observation
```{r}
test.ratings <- movie.by.user[which(movie.by.user$movieId %in% c(150,296,380,590)),c(2,3)]
ratings.file[which(ratings.file$userId == 191 & ratings.file$movieId %in% c(150,296,380,590)),3] <- "NA"

```

### a Prediction using user-user similarity:

```{r}
ratings.file <- data.frame(lapply(ratings.file, as.character),stringsAsFactors = FALSE)
neighbours_191 <- c("225","317","415","50","568") 
jacccard_191 <- c(0.2420382,0.4033613,0.3255814,0.2009804,0.2105263)
ros <- c("191",neighbours_191)

user225 <- subset(ratings.file,ratings.file$userId==225)
user317 <- subset(ratings.file,ratings.file$userId==317)
user415 <-subset(ratings.file,ratings.file$userId==415)
user50 <- subset(ratings.file,ratings.file$userId==50)
user568 <-subset(ratings.file,ratings.file$userId==568)

utility <- matrix(as.numeric(0),nrow = 6,ncol = 29)

movie_list <- intersect(movie.by.user$movieId,user225$movieId)
movie_list <- c(unique(movie_list),intersect(movie.by.user$movieId,user415$movieId))
movie_list <- c(unique(movie_list),intersect(movie.by.user$movieId,user50$movieId))
movie_list <- c(unique(movie_list),intersect(movie.by.user$movieId,user568$movieId))
movie_list <- c(unique(movie_list),intersect(movie.by.user$movieId,user317$movieId))
movie_list <- unique(movie_list)

colnames(utility) <- movie_list
rownames(utility) <- ros

for(i in 1:length(ros)){
  subset_data <- subset(ratings.file,ratings.file$userId==ros[i])
  temp.user.id <- subset_data[1,1]
  for(j in 1:length(movie_list)){
    movie.id <- movie_list[j]
    rating.value <- subset_data[which(movie.id  == subset_data$movieId),3]
    utility[temp.user.id,movie.id] <- ifelse(identical(rating.value, character(0)),NA ,rating.value)
  }
}

utility


```

### Predicition
```{r}
random_neighbours <- neighbours_191[sample(1:length(neighbours_191), 3, replace = FALSE)]
final_ans <- data.frame()
rmse.n <- 0
for(i in 1:nrow(test.ratings)){
  numerator <- 0
  denominator <- 0
  for(j in 1:length(random_neighbours)){
    neighbour.id<- random_neighbours[j]
    jaccard.sim <- as.numeric(jacccard_191[which(neighbours_191 == neighbour.id)])
    index <- which(ros == neighbour.id)
    movie.index<- which(movie_list == test.ratings$movieId[i] )
    
    rating.value <- as.numeric(utility[index, movie.index])
    numerator <-  numerator + (jaccard.sim * rating.value)
    denominator <- denominator + jaccard.sim
  }
  movieId = as.character(test.ratings$movieId[i])
  ratings = (numerator) / (denominator)
  rmse.n = round((ratings - test.ratings$rating[i])^2,4)
  final_ans  <- rbind (final_ans, data.frame(movieId, ratings))
  
}

movieId = "RMSE"
ratings <- sqrt(rmse.n/4)

final_ans  <- rbind (final_ans, data.frame(movieId, ratings))
cat("\n User ID 191, 5 random user IDs: ", ros)
cat("\n Using user-user similarity, User ID 191 will rate the movies as follows:")
print(final_ans)
```



### item-item

```{r}
transpose_utility <- matrix(as.numeric(0),nrow = 29,ncol = 6)
transpose_utility <- t(utility)
transpose_utility

```
```{r}
mean_utility <- transpose_utility
mean_utility[is.na(mean_utility)] <- 0

for(i in 1:nrow(transpose_utility)){
  temp.val <- mean_utility[i,]
  temp.mean <- mean(as.double(temp.val),na.rm = T)
  mean_utility[i,] <- round(temp.mean - as.double(mean_utility[i,]),3)
}

colnames(mean_utility) <- ros
rownames(mean_utility) <- movie_list
cat("\n\n")
mean_utility
```
### new cosine similarity

```{r}
my.cosine <- function(x, y) {
  sum(x*y)/(norm(x, type="2") * norm(y, type="2"))
}
non_mean_utility <- transpose_utility

non_mean_utility[is.na(non_mean_utility)] <- 0
non_mean_utility[which(non_mean_utility == "NA"),1] <- 0
new_cosine <- array(dim = c(nrow(non_mean_utility),4))

for(j in 1:nrow(test.ratings)){
  for(i in 1:nrow(non_mean_utility)){
    if(movie_list[i] %in% test.ratings$movieId){
      new_cosine[i,j] <- 1
    } else{
      colInMovieList <- which(movie_list == test.ratings[j,1])
      new_cosine[i,j] <- round(my.cosine(as.double(non_mean_utility[i,]), as.double(non_mean_utility[colInMovieList,])),3)
    }
  }
}
row.names(new_cosine) <- movie_list
colnames(new_cosine) <- test.ratings$movieId
new_cosine
```

### prediticion item item
```{r}

transpose_utility[is.na(transpose_utility)] <- 0
transpose_utility[which(transpose_utility == "NA"),1] <- 0

item_final_ans <- data.frame()
rmse.n = 0
neighbours_movies <- movie_list[which(!(movie_list %in% test.ratings$movieId))]
random_neighbours <- neighbours_movies[sample(1:length(neighbours_movies), 3, replace = FALSE)]

for(i in 1:nrow(test.ratings)){
  k=0
  movieId <- test.ratings[i,]$movieId
  numerator <- 0
  denominator <- 0
  user_cosine_values <- new_cosine[,i]
  for(j in 1:length(random_neighbours)){
    
    movieNeighbourIdIndex <- which(random_neighbours[j] == movie_list)
    if(user_cosine_values[[movieNeighbourIdIndex]] > 0  && user_cosine_values[[movieNeighbourIdIndex]] < 1){
      numerator <-  numerator + (user_cosine_values[[movieNeighbourIdIndex]] * as.numeric(transpose_utility[movieNeighbourIdIndex,1]))
      denominator <- denominator + user_cosine_values[[movieNeighbourIdIndex]]
    }
  }
  ratings = round((numerator) / (denominator),4)
  rmse.n = rmse.n + round((ratings - test.ratings$rating[i])^2,4)
  item_final_ans  <- rbind (item_final_ans, data.frame(movieId, ratings))
}

movieId = "RMSE"
ratings <- sqrt(rmse.n/4)

item_final_ans  <- rbind (item_final_ans, data.frame(movieId, ratings))

cat("\n User ID 191, 5 random user IDs: ", ros)
cat("\n Using user-user similarity, User ID 191 will rate the movies as follows:")
print(item_final_ans)
```












